---
title: "NSW data exploration"
author: "David Anderson z5343521"
date: "19/03/2022"
output: html_document
---

I want to explore the energy demand dataset for NSW and see if I can find any interesting trends in the demand over time or demand in relation to temperature which could be used to create a valid forecast model.
I will be using a set of data that has already been cleaned.

```{r}
# Import libraries
library(lubridate)
library(ggplot2)
library(dplyr)
library(hms)
library(Metrics)

# Import cleaned data
NSW_df <- read.csv("Cleaned_Data_w_fcst.csv")
NSW_df$DATETIME <- ymd_hms(NSW_df$DATETIME)
```

It's unlikely to give us any results but let's look at the relation between temp and demand per year to see if there is magically a useful correlation.

```{r fig.height = 10, fig.width = 30}
# Plot the temp and demand data per year
ggplot(NSW_df, aes(x = TEMPERATURE, y = TOTALDEMAND)) +
  geom_point() +
  facet_wrap(~year(DATETIME))
```

We can see a recurring 'U' shaped pattern that is very similar year to year. A simple regression model would have far too much error based on how widely spread the data points are from an imagined line of best fit.
Let's see if we can find some trends and/or make a linear regression model work somehow, even though it would be wildly inaccurate.

Let's look at the trend in demand and temperature over time individually.

```{r fig.height = 10, fig.width = 30}
# Plot the temp and demand data on a dual y axis against time to see any correlation.
ggplot(NSW_df, aes(x = DATETIME)) +
  geom_line(aes(y = TEMPERATURE), colour = 'red') + 
  geom_line(aes(y = TOTALDEMAND/400), colour = 'blue') +
  scale_y_continuous(name = 'Temperature', sec.axis = sec_axis(~.*400, name = 'Demand'))

```

This does show some trends:
      - There is definitely a 'W' shape in the energy usage every year (and a more emphasized one in Temp). This likely corresponds to high energy usage for cooling and heating in summer and winter, with lower usage in the transitory months.
      - Energy demand in trending slightly downward over the years.
      - Is temperature trending slightly upwards over the years? Yes but very slowly, and not consistently.

It would be more informative to see this graph on a per year scale I think.

```{r fig.height = 10, fig.width = 30}
# Plot the temp and demand data per year
ggplot(NSW_df, aes(x = DATETIME)) +
  geom_line(aes(y = TEMPERATURE), colour = 'red') + 
  geom_line(aes(y = TOTALDEMAND/250), colour = 'blue') +
  scale_y_continuous(name = 'Temperature', sec.axis = sec_axis(~.*250, name = 'Demand')) +
  facet_wrap(~year(DATETIME), scales = 'free')
  
```

There is definitely a correlation between temperature and demand but we can see that both are quite chaotic and that the correlation is not strongly followed all of the time.
All of the years exhibit similar trends. Note that the demand data is scaled by a factor of 250 and the y-axis scales are free to fit the data, so we will not notice any increase/reduction over time in relative temp/demand in this plot.

The data seems to show that there is more chaotic energy usage in the months of Nov - Feb. Can we verify this with a plot?

```{r fig.height = 10, fig.width = 30}
# Boxplot temp per month
ggplot(NSW_df, aes(x = month(DATETIME), y = TEMPERATURE, group = month(DATETIME))) + geom_boxplot() + xlim(0, 13)

# Boxplot demand per month
ggplot(NSW_df, aes(x = month(DATETIME), y = TOTALDEMAND, group = month(DATETIME))) + geom_boxplot() + xlim(0, 13)
```

There is definitely more chaotic temperatures and demand between Nov - Feb (outliers in the boxplots). These dates do not have a much higher average energy usage however.

I'd love to see if there was an interesting correlation between demand and temperature if the temperature was measured in total degrees away from the yearly average.

```{r fig.height = 10, fig.width = 30}
# Calculate the absolute value of the difference between each temperature and the average temperature for that year.
NSW_df <- NSW_df %>%
  mutate(Year = year(DATETIME)) %>%
  group_by(Year) %>%
  mutate(Temp_Fluctuation = abs(TEMPERATURE - mean(TEMPERATURE)))

# Plot
ggplot(NSW_df, aes(x = DATETIME)) +
  geom_line(aes(y = Temp_Fluctuation), colour = 'red') + 
  geom_line(aes(y = TOTALDEMAND/250), colour = 'blue') +
  scale_y_continuous(name = 'Temperature', sec.axis = sec_axis(~.*250, name = 'Demand')) +
  facet_wrap(~year(DATETIME), scales = 'free')
```

This looks like a pretty strong correlation..... The demand is again scaled by 250 to better see the correlation.

Let's look at what the demand vs temp fluctuation scatterplot looks like.

```{r fig.height = 10, fig.width = 30}
# Plot demand against temperature fluctuation per year
ggplot(NSW_df, aes(x = Temp_Fluctuation, y = TOTALDEMAND)) +
  geom_point() +
  geom_smooth(method = lm) +
  facet_wrap(~year(DATETIME))
```

The 'U' shape is gone.

Let's make some linear models, for fun.

```{r}
# First linear model between temp and demand
lm1 <- lm(TOTALDEMAND ~ TEMPERATURE, data = NSW_df)
summary(lm1)
plot(lm1)

NSW_df$lm1_fcst <- predict(lm1)
NSW_df$lm1_res <- NSW_df$TOTALDEMAND - NSW_df$lm1_fcst
```

An R squared of 0.02 is useless. The residuals vs fitted is the shape that we expect. Normal Q-Q shows that our model does not capture extreme values well at all.

```{r}
# A linear model between temp fluctuation away from the yearly average will likely fix the Residuals vs fitted graph
lm2 <- lm(TOTALDEMAND ~ Temp_Fluctuation, data = NSW_df)
summary(lm2)
plot(lm2)

NSW_df$lm2_fcst <- predict(lm2)
NSW_df$lm2_res <- NSW_df$TOTALDEMAND - NSW_df$lm2_fcst
```

An R squared of 0.08 is still useless. The Residuals vs Fitted was 'better'. Normal Q-Q shows that we capture some extreme values a bit better. The model is still useless overall.

Let's compare this simple model to the forecast models given in the data files.

```{r}
## Boxplots
#boxplot(NSW_df$Diff1, NSW_df$Diff32, NSW_df$lm1_res, NSW_df$lm2_res)

# Caluculate five number summaries
fivenum(NSW_df$Fcst1_res)
fivenum(NSW_df$Fcst32_res)
fivenum(NSW_df$lm1_res)
fivenum(NSW_df$lm2_res)

# Calculate RSME
print(" ")
print("RSME")
sqrt(mean(NSW_df$Fcst1_res^2))
sqrt(mean(NSW_df$Fcst32_res^2))
sqrt(mean((NSW_df$lm1_res)^2))
sqrt(mean((NSW_df$lm2_res)^2))

# Calculate MAE
print(" ")
print("MAE")
mae(NSW_df$TOTALDEMAND, NSW_df$Fcst1)
mae(NSW_df$TOTALDEMAND, NSW_df$Fcst32)
mae(NSW_df$TOTALDEMAND, NSW_df$lm1_fcst)
mae(NSW_df$TOTALDEMAND, NSW_df$lm2_fcst)

# Calculate MAD
print(" ")
print("MAD")
mad(NSW_df$TOTALDEMAND - NSW_df$Fcst1)
mad(NSW_df$TOTALDEMAND - NSW_df$Fcst32)
mad(NSW_df$TOTALDEMAND - NSW_df$lm1_fcst)
mad(NSW_df$TOTALDEMAND - NSW_df$lm2_fcst)

```

These metrics show that our models perform terribly compared to the other provided.

It's pretty unlikely to be able to improve a linear model to a point where it is useful.

We have identified that a ML model is likely going to be our best option for this project. 
Let's see if we can identify any factors that could be used to help train a ML model.

```{r}
# Plot demand over different time intervals to see any trends
ggplot(NSW_df, aes(x = year(DATETIME), y = TOTALDEMAND, group = year(DATETIME))) + geom_boxplot()
ggplot(NSW_df, aes(x = week(DATETIME), y = TOTALDEMAND, group = week(DATETIME))) + geom_boxplot()
ggplot(NSW_df, aes(x = wday(DATETIME), y = TOTALDEMAND, group = wday(DATETIME))) + geom_boxplot()
ggplot(NSW_df, aes(x = as_hms(DATETIME), y = TOTALDEMAND, group = as_hms(DATETIME))) + geom_boxplot()

```

We have already established that demand is decreasing over the years. This is likely due to more energy efficient appliances and practices.
Demand over the week is just showing the more extreme values in summer months and a 'W' shape previously identified due to more energy being used to cool/heat in summer/winter and less so in transitory months.
It looks like energy demand is down over the weekend, likely due to commercial and industrial requirements being lower.
The average and spread of energy demand increases over time of the day from around 4:30am and starts decreasing at around 5:30pm, likely due to commercial and industrial requirements again. As the average usage increases, so does the spread.

The key variables that we can identify from this for use in a ML model would be:
    - Weekday vs weekend
    - On-peak vs off-peak energy times during the day
    - Year
    - Season


```{r}
# Export the clean_data_w_fcst to csv
write.csv(NSW_df, "Explored_data.csv", row.names = FALSE)
```

What about using demand from 30, 60 and 90 mins prior as a predictor?


```{r}
NSW_df$dem_30 <- NSW_df$TOTALDEMAND
NSW_df <- NSW_df %>%
  mutate(dem_30 = lag(dem_30, 1)) %>%
  na.omit()

NSW_df$dem_60 <- NSW_df$TOTALDEMAND
NSW_df <- NSW_df %>%
  mutate(dem_60 = lag(dem_60, 2)) %>%
  na.omit()

NSW_df$dem_90 <- NSW_df$TOTALDEMAND
NSW_df <- NSW_df %>%
  mutate(dem_90 = lag(dem_90, 3)) %>%
  na.omit()
```






```{r}
ggplot(NSW_df, aes(x = dem_30, y = TOTALDEMAND)) + geom_point() #, group = year(DATETIME)))
ggplot(NSW_df, aes(x = dem_60, y = TOTALDEMAND)) + geom_point() #, group = year(DATETIME)))
ggplot(NSW_df, aes(x = dem_90, y = TOTALDEMAND)) + geom_point() #, group = year(DATETIME)))
```




```{r}
# 
lmd30 <- lm(TOTALDEMAND ~ dem_30, data = NSW_df)
summary(lmd30)
plot(lmd30)

```






```{r}
NSW_df$temp_30 <- NSW_df$TEMPERATURE
NSW_df <- NSW_df %>%
  mutate(temp_30 = lag(temp_30, 1)) %>%
  na.omit()
```






```{r}
ggplot(NSW_df, aes(x = temp_30, y = TOTALDEMAND)) + geom_point()
```




```{r}
# 
lmt30 <- lm(TOTALDEMAND ~ temp_30, data = NSW_df)
summary(lmt30)
plot(lmt30)

```


```{r}
M <- cor(NSW_df)
corrplot(M, method = "number")
```


