---
title: "Feature_Selection"
author: "David Anderson"
date: "4/3/2022"
output: html_document
---

Want to see if we can come up with some mroe useful features than just temperature. We have seen that temperature by itself is a pretty bad indicator of demand.

```{r fig.height = 10, fig.width = 30}
# Import libraries
library(lubridate)
library(ggplot2)
library(dplyr)
library(hms)
library(Metrics)
library(corrplot)

# Import cleaned data
features <- read.csv("Cleaned_Data_mkII.csv")
features$DATETIME <- ymd_hms(features$DATETIME)

#Look at the relation between temp and demand per year to show the correlation.
ggplot(features, aes(x = TEMPERATURE, y = TOTALDEMAND)) +
  geom_point() +
  facet_wrap(~year(DATETIME))
```

What about using demand from 30, 60 and 90 mins prior as a predictor?

```{r}
features$dem_30 <- features$TOTALDEMAND
features <- features %>%
  mutate(dem_30 = lag(dem_30, 1)) %>%
  na.omit()

features$dem_60 <- features$TOTALDEMAND
features <- features %>%
  mutate(dem_60 = lag(dem_60, 2)) %>%
  na.omit()

features$dem_90 <- features$TOTALDEMAND
features <- features %>%
  mutate(dem_90 = lag(dem_90, 3)) %>%
  na.omit()

ggplot(features, aes(x = dem_30, y = TOTALDEMAND)) + geom_point()
ggplot(features, aes(x = dem_60, y = TOTALDEMAND)) + geom_point()
ggplot(features, aes(x = dem_90, y = TOTALDEMAND)) + geom_point()
```

These are obviously much better predictors. How well can we predict current demand from demand 30 mins ago, using a simple linear model?

```{r}
# 
lmd30 <- lm(TOTALDEMAND ~ dem_30, data = features)
summary(lmd30)
plot(lmd30)

```

Not bad.... Let's explore some other potential features using a coorplot.

```{r}
# Import cleaned data
weather <- read.csv("weather.csv")
#features$DATETIME <- ymd_hms(features$DATETIME)

features$date <- date(features$DATETIME)
weather$date <- as.Date(with(weather, paste(Year, Month, Day, sep="-")), "%Y-%m-%d")

weather <- weather[, -c(1, 2, 3, 4, 5, 7, 8)]

# Do a left merge on dem and temp data
features <- left_join(features, weather, by = 'date')
```




We have seen that weekday, month and time are a rough indicator of change, but not a very good one.

```{r}
features$day <- wday(features$DATETIME)
features$month <- month(features$DATETIME)
features$time <- as_hms(features$DATETIME)
```






If we can use time from 30, 60 and/or 90 mins ago as useful features, it would mean that we can use historic temp data in a model, rather than forecasted temp data. This would be more accurate as an input but will it be as useful as a forecast temp value?

```{r}
features$temp_30 <- features$TEMPERATURE
features <- features %>%
  mutate(temp_30 = lag(temp_30, 1)) %>%
  na.omit()

features$temp_60 <- features$TEMPERATURE
features <- features %>%
  mutate(temp_60 = lag(temp_60, 2)) %>%
  na.omit()

features$temp_90 <- features$TEMPERATURE
features <- features %>%
  mutate(temp_90 = lag(temp_90, 3)) %>%
  na.omit()
```





```{r}
features$time <- (hour(features$time)*60 + minute(features$time))/30
features <- features[, -c(3, 7)]
```






```{r fig.height = 10, fig.width = 30}
M <- cor(features)
corrplot(M, method = "number")
```


```{r}
fsmlm <- lm(TOTALDEMAND ~ ., data = features)
summary(fsmlm)
plot(fsmlm)

print(paste("MAE:", mae(features$TOTALDEMAND, predict(fsmlm))))

print(paste("MAD:", mad(features$TOTALDEMAND - predict(fsmlm))))
```

```{r}
fsmlm2 <- lm(TOTALDEMAND ~ temp_30 + dem_30 + dem_60, data = features)
summary(fsmlm2)
plot(fsmlm2)


print(paste("MAE:", mae(features$TOTALDEMAND, predict(fsmlm2))))

print(paste("MAD:", mad(features$TOTALDEMAND - predict(fsmlm2))))
```


```{r}
plot(features$TOTALDEMAND, predict(fsmlm2))

```




```{r}
## Export the clean_data_w_fcst to csv
#write.csv(features, "Explored_data.csv", row.names = FALSE)
```


```{r}
features$change_in_demand <- features$TOTALDEMAND - features$dem_30

```




```{r}
fslm <- lm(change_in_demand ~ TEMPERATURE, data = features)
summary(fslm)
plot(fslm)


print(paste("MAE:", mae(features$TOTALDEMAND, predict(fslm))))

print(paste("MAD:", mad(features$TOTALDEMAND - predict(fslm))))
```























