---
title: "NSW data exploration"
authors: "Abdul	El-Hamawi	z5019165, Chris	Strods	z5329477, David Anderson z5343521, Jamie	Twiss	z5353394, Shubhankar	Dutta	z5304573, Sonal	Chawla	z5092985"
date: "19/03/2022"
output: html_document
---

We want to explore the energy demand dataset for NSW and highlight interesting trends.
We will be using a set of data that has already been cleaned.

```{r}
# Import libraries
library(lubridate)
library(ggplot2)
library(dplyr)
library(hms)
library(Metrics)
library(corrplot)

# Import cleaned data
Clean_df <- read.csv("Cleaned_Data.csv")
Clean_df$DATETIME <- ymd_hms(Clean_df$DATETIME)
```
Graph temperature vs demand over all of the data

```{r fig.height = 10, fig.width = 15}
# Plot the temp and demand data per year
ggplot(Clean_df, aes(x = TEMPERATURE, y = TOTALDEMAND)) +
  ggtitle("Temperature vs Demand for 2010-2021") +
  xlab('Temperature (deg celcius)') +
  ylab('Demand (MW)') +
  geom_point()
```

Demand is likely to fluctuate over the time of the day.

```{r fig.height = 10, fig.width = 15}
# Boxplot demand vs time of day
ggplot(Clean_df, aes(x = as_hms(DATETIME), y = TOTALDEMAND, group = as_hms(DATETIME))) +
  ggtitle("Demand vs Time of Day") +
  xlab('Time') +
  ylab('Demand (MW)') +
  geom_boxplot()
```
Day of the week is less likely to be an indicator but we will look to see

```{r fig.height = 10, fig.width = 15}
# Boxplot demand over day of the week
ggplot(Clean_df, aes(x = wday(DATETIME), y = TOTALDEMAND, group = wday(DATETIME))) +
  ggtitle("Demand vs Day of Week") +
  xlab('Day of Week') +
  ylab('Demand (MW)') +
  geom_boxplot()
```

Does week of the year have an impact?

```{r fig.height = 10, fig.width = 15}
# Boxplotdemand vs week of the year
ggplot(Clean_df, aes(x = week(DATETIME), y = TOTALDEMAND, group = week(DATETIME))) +
  ggtitle("Demand vs Week of Year") +
  xlab('Week of Year') +
  ylab('Demand (MW)') +
  geom_boxplot()
```
Month of the year will show the same trend at a higher level

```{r fig.height = 10, fig.width = 15}
# Boxplot demand vs month of year
ggplot(Clean_df, aes(x = month(DATETIME), y = TOTALDEMAND, group = month(DATETIME))) +
  ggtitle("Demand vs Month of Year") +
  xlab('Month of Year') +
  ylab('Demand (MW)') +
  geom_boxplot()
```

Finally, let's see what the correlation is between demand, temperature, day, month and time of day.

```{r fig.height = 10, fig.width = 15}
# Convert column names to something nicer
Clean_df$Demand <- Clean_df$TOTALDEMAND
Clean_df$Temperature <- Clean_df$TEMPERATURE
Clean_df$Day <- wday(Clean_df$DATETIME)
Clean_df$Month <- month(Clean_df$DATETIME)
Clean_df$Time <- as_hms(Clean_df$DATETIME)
Clean_df <- Clean_df[, -c(1,2,3)]

# Convert time to numeric
Clean_df$Time <- (hour(Clean_df$Time)*60 + minute(Clean_df$Time))/30

# Plot corrplot
M <- cor(Clean_df)
corrplot(M, method = "number")
```


