---
title: "Capstone_project_Group_E"
authors: "Abdul	El-Hamawi	z5019165, Chris	Strods	z5329477, David Anderson z5343521, Jamie	Twiss	z5353394, Shubhankar	Dutta	z5304573, Sonal	Chawla	z5092985"
date: "12/03/2022"
output: html_document
---



```{r}
# Import libraries
library(lubridate)
library(car)
library(ggplot2)
library(tidyverse)
library(dplyr)

# Import temperature and demand datasets       
temp_nsw <- read.csv(unz('temperature_nsw.csv.zip', 'temperature_nsw.csv'), header = T)
dem_nsw <- read.csv(unz('totaldemand_nsw.csv.zip', 'totaldemand_nsw.csv'), header = T)

# Get rid of columns that we don't require
temp_nsw <- temp_nsw[-c(1)]
dem_nsw <- dem_nsw[-c(3)]

# Split DATETIME into Date and Time columns
dem_nsw <- dem_nsw %>%
  mutate("Date" = str_extract(DATETIME, "[0-9]?[0-9]/[0-9]?[0-9]/[0-9]{4}"))
dem_nsw <- dem_nsw %>%
  mutate("Time" = str_extract(DATETIME, "[0-9]?[0-9]:[0-9]{2}"))
temp_nsw <- temp_nsw %>%
  mutate("Date" = str_extract(DATETIME, "[0-9]?[0-9]/[0-9]?[0-9]/[0-9]{4}"))
temp_nsw <- temp_nsw %>%
  mutate("Time" = str_extract(DATETIME, "[0-9]?[0-9]:[0-9]{2}"))

# Delete DATETIME columns
temp_nsw <- temp_nsw[-c(1)]
dem_nsw <- dem_nsw[-c(1)]

# Check for any missing values.
temp_nsw[rowSums(is.na(temp_nsw)) > 0, ] 
dem_nsw[rowSums(is.na(dem_nsw)) > 0, ]

```
No missing values.

Our data is from 01-01-2010 until 18-03-2021 which is 4094 days.

Check to see how many of each unique time value our dataframes contain.

```{r}
# Check the count of all unique time values in each data frame
as.data.frame(table(temp_nsw$Time))
as.data.frame(table(dem_nsw$Time))
```

This lines up with what we would expect in dem_nsw, if this follows the correct pattern of sampling every 30 mins across the span of our data we can use it as the data to left_join temp_nsw onto and resolve any issues from there.

We can also check that each date in the data contains 48 samples

```{r}
filter(as.data.frame(table(dem_nsw$Date)), Freq != 48)
```

This is as expected, the last date (18/3/2021) Was only sampled once.

```{r}
# Remove the last row to make this easier
dem_nsw <- dem_nsw[-196513, ]

# Check for duplicates
dem_nsw[duplicated(dem_nsw[, ]), ]
```

Since we have 4095 samples of each timestamp, every date occurs 48 times and there are no duplicated records, we can conclude that our data follows the time sampling format we require.

Thus we have all of the times that we wish to include from 01-01-2010 until 18-03-2021, occurring in 30 minute intervals. 
We can do a left merge onto this data in order to preserve it's structure and finish cleaning the temp data in this new structure.

```{r}
# Do a left merge on dem and temp data
cleaned_temp_dem_df <- left_join(dem_nsw, temp_nsw, by = c('Date', 'Time'))
```

Why does this give extra rows?

```{r}
# Check where the extra rows have been inserted
filter(as.data.frame(table(cleaned_temp_dem_df$Date)), Freq != 48)
```

The extra rows are all just duplicated rows. Why did they appear though??

```{r}
# Get rid of duplicates
cleaned_temp_dem_df <- cleaned_temp_dem_df[!duplicated(cleaned_temp_dem_df[, ]), ]
```

Our data is now in the correct time series format. We just need to make sure that the temperature data is all there and we can move on to visualization. Incorrectly entered data may still be an issue, this shoudl appear in visualization.

```{r}
# Check for NaN values
cleaned_temp_dem_df[rowSums(is.na(cleaned_temp_dem_df)) > 0, ]
```

Fill NaN values.

```{r}
# Replcae any NaN temperature data with the value before it. I would like this to be the average of the temperature before it and the one after it instead. Maybe re-write this later
cleaned_temp_dem_df <- cleaned_temp_dem_df %>%
  mutate(TEMPERATURE = replace(TEMPERATURE, is.nan(TEMPERATURE), NA)) %>%
  fill(TEMPERATURE)
```

```{r}
# Convert date and time back to a datetime
cleaned_temp_dem_df$DATETIME <- paste(cleaned_temp_dem_df$Date, cleaned_temp_dem_df$Time)
cleaned_temp_dem_df$DATETIME <- dmy_hm(cleaned_temp_dem_df$DATETIME)
cleaned_temp_dem_df <- cleaned_temp_dem_df[, -c(2,3)]
```

This concludes the data cleaning for now I think. More to come after visualization or as people proof this code, I'm sure.

```{r}
# Export the cleaned_temp_dem_df to csv
write.csv(cleaned_temp_dem_df, "Cleaned_Data.csv", row.names = FALSE)
```

